---
title: "Finding Alpha in Chicago's Disinvested Communities"
subtitle: "A Spatial-Alpha Index Combining 2SFCA Transit Accessibility and Local Moran's I"
author: "Zhen Zeng"
date: today
output-file: index.html
format:
  html:
    toc: true
    toc-float: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    embed-resources: true
    theme: cosmo
    fig-width: 9
    fig-height: 7
    smooth-scroll: true
execute:
  warning: false
  message: false
  results: hide
knitr:
  opts_chunk:
    fig.align: center
---

```{r one-time-setup}
#| label: one-time-setup
#| eval: false
#| include: true

# One-time install of packages missing from this machine's R library.
# (Detected via installed.packages(); already-installed deps are not relisted.)
install.packages(c("here", "rgeoda"))

# Optional: pin versions with renv
# renv::init()
# renv::snapshot()
```

```{r setup}
#| label: setup
#| include: false

# ---- Reproducibility ----
# Project can be pinned with {renv}; {here} keeps paths relative to the repo.

library(here)
library(tidyverse)
library(sf)
library(tigris)        # Census TIGER/Line geographies
library(tidycensus)    # ACS pull (requires CENSUS_API_KEY env var)
library(rgeoda)        # Local Moran's I (LISA)
library(leaflet)
library(gt)
library(DT)
library(crosstalk)
library(htmltools)
library(httr2)
library(janitor)
library(scales)
library(RColorBrewer)

# tigris config: cache shapefiles, return sf, use 2022 vintage
options(
  tigris_use_cache = TRUE,
  tigris_class     = "sf"
)

# Census API key from environment (never hardcode)
if (Sys.getenv("CENSUS_API_KEY") != "") {
  census_api_key(Sys.getenv("CENSUS_API_KEY"), install = FALSE)
}

# Projected CRS for Chicago: NAD83 / Illinois East (meters)
CRS_PROJ <- 26971
CRS_GEO  <- 4326

# Year vintage for ACS + permit window
ACS_YEAR     <- 2022
PERMIT_YEARS <- 2019:2024

# 2SFCA catchment radius (meters)
CATCHMENT_M <- 1000
```

# 1. Introduction & Research Question

::: {.callout-note title="Project links"}
This report is published at [uchizhen.github.io](https://uchizhen.github.io/).
The full project repository is available at
[github.com/UChiZhen/UChiZhen.github.io](https://github.com/UChiZhen/UChiZhen.github.io).
:::

Impact investors increasingly seek **place-based alpha** — neighborhoods whose
underlying fundamentals (transit, density, demographics) suggest stronger
performance than current capital flows imply. This report operationalizes that
intuition for **Chicago at the Census tract level** by combining:

- a **2-Step Floating Catchment Area (2SFCA)** measure of accessibility to the
  CTA "L" rapid-transit network (the *high-potential* signal), and
- **Local Moran's I (LISA)** on a composite **Spatial Alpha Index** that
  subtracts realized investment intensity (building-permit dollars per housing
  unit) from that accessibility score.

> **Research question.** *Where in Chicago do statistically significant clusters
> of high transit accessibility coincide with low realized investment, and
> what do those clusters look like demographically?*

The **High–High LISA clusters** of the alpha index are the report's headline
output: contiguous areas of "transit-rich but capital-starved" tracts that
constitute the strongest *prima facie* targets for catalytic capital.

# 2. Data Sources & Cleaning

The analysis combines three independent, free sources rather than relying solely on ACS:

| # | Source | Role | Endpoint |
|---|--------|------|----------|
| 1 | **Chicago Data Portal — Building Permits** | Investment intensity (low = under-invested) | `data.cityofchicago.org/resource/ydr8-5enu` |
| 2 | **Chicago Data Portal — CTA "L" Stations** | Supply side of 2SFCA | `data.cityofchicago.org/resource/8pix-ypme` |
| 3 | **ACS 5-yr (`tidycensus`)** | Population & housing denominators, descriptive covariates | Census Bureau API |

::: {.panel-tabset}

## Geographies

```{r geographies}
#| label: geographies
#| cache: true

# Cook County tracts (ACS 2022 vintage)
cook_tracts <- tracts(
  state = "IL", county = "Cook",
  cb = TRUE, year = ACS_YEAR
) |>
  st_transform(CRS_PROJ) |>
  select(GEOID, NAME_tract = NAME, ALAND)

# City of Chicago boundary (TIGER Place file)
chicago_boundary <- places(state = "IL", cb = TRUE, year = ACS_YEAR) |>
  filter(NAME == "Chicago") |>
  st_transform(CRS_PROJ) |>
  st_union()

# Clip tracts to city: keep tracts whose centroid lies in Chicago.
# (Centroid test avoids slivers from boundary irregularities.)
tract_centroids <- st_centroid(cook_tracts)
in_city <- st_within(tract_centroids, chicago_boundary, sparse = FALSE)[, 1]
chi_tracts <- cook_tracts[in_city, ]

# Drop O'Hare-style non-residential outliers later via housing-unit filter.
```

## ACS pull

```{r acs}
#| label: acs
#| cache: true

acs_raw <- get_acs(
  geography = "tract",
  state     = "IL",
  county    = "Cook",
  variables = c(
    total_pop     = "B01003_001",
    housing_units = "B25001_001",
    med_hh_income = "B19013_001",
    renter_occ    = "B25003_003",
    occ_total     = "B25003_001",
    white_nh      = "B03002_003",
    total_race    = "B03002_001"
  ),
  year      = ACS_YEAR,
  survey    = "acs5",
  output    = "wide",
  geometry  = FALSE
) |>
  clean_names() |>
  transmute(
    geoid          = geoid,
    total_pop      = total_pop_e,
    housing_units  = housing_units_e,
    med_hh_income  = med_hh_income_e,
    pct_renter     = ifelse(occ_total_e > 0, renter_occ_e / occ_total_e, NA_real_),
    pct_nonwhite   = ifelse(total_race_e > 0, 1 - white_nh_e / total_race_e, NA_real_)
  )
```

## CTA "L" stations

```{r lstops}
#| label: lstops
#| cache: true

l_stops_path <- here("data", "raw", "cta_l_stops.geojson")

if (!file.exists(l_stops_path)) {
  # Socrata GeoJSON endpoint — no auth needed for moderate pulls
  req <- request("https://data.cityofchicago.org/resource/8pix-ypme.geojson") |>
    req_url_query(`$limit` = 5000)
  resp <- req_perform(req)
  writeBin(resp_body_raw(resp), l_stops_path)
}

l_stops_raw <- st_read(l_stops_path, quiet = TRUE) |>
  st_transform(CRS_PROJ)

# Each row is a station ENTRANCE; collapse to one point per station_id
l_stations <- l_stops_raw |>
  group_by(station_name) |>
  summarise(geometry = st_centroid(st_union(geometry)), .groups = "drop") |>
  mutate(station_id = row_number())
```

## Building permits (investment proxy)

```{r permits}
#| label: permits
#| cache: true

permits_path <- here("data", "raw", "chi_permits.csv")

if (!file.exists(permits_path)) {
  # Pull only what we need: residential-relevant permits 2019-2024, with $ valuation
  # Dataset: ydr8-5enu (Building Permits)
  # We page through Socrata to stay polite (50k rows per pull)
  base <- "https://data.cityofchicago.org/resource/ydr8-5enu.csv"
  fields <- "id,permit_,permit_type,application_start_date,issue_date,reported_cost,latitude,longitude"
  where  <- sprintf(
    "issue_date >= '%d-01-01T00:00:00' AND issue_date < '%d-01-01T00:00:00' AND reported_cost > 0 AND latitude IS NOT NULL",
    min(PERMIT_YEARS), max(PERMIT_YEARS) + 1
  )

  pulls <- list()
  offset <- 0
  repeat {
    req <- request(base) |>
      req_url_query(
        `$select` = fields,
        `$where`  = where,
        `$limit`  = 50000,
        `$offset` = offset
      )
    # Force all-character on read so column types are consistent across chunks
    # (Socrata sometimes returns numeric-looking IDs that later contain letters,
    # which makes readr guess different types per chunk -> bind_rows fails.)
    chunk <- read_csv(
      resp_body_string(req_perform(req)),
      show_col_types = FALSE,
      col_types      = cols(.default = "c")
    )
    if (nrow(chunk) == 0) break
    pulls[[length(pulls) + 1]] <- chunk
    if (nrow(chunk) < 50000) break
    offset <- offset + 50000
  }
  permits_dl <- bind_rows(pulls)
  write_csv(permits_dl, permits_path)
}

# Read cached file as character, coerce numerics explicitly below.
permits_raw <- read_csv(
  permits_path,
  show_col_types = FALSE,
  col_types      = cols(.default = "c")
)

# Clean: keep construction-related permits, drop fee-only / sign permits
permits_clean <- permits_raw |>
  clean_names() |>
  mutate(
    reported_cost = suppressWarnings(as.numeric(reported_cost)),
    latitude      = suppressWarnings(as.numeric(latitude)),
    longitude     = suppressWarnings(as.numeric(longitude))
  ) |>
  filter(
    !is.na(latitude), !is.na(longitude),
    !is.na(reported_cost), reported_cost > 0,
    str_detect(
      str_to_lower(permit_type),
      "new construction|renovation|alteration|addition"
    )
  )

permits_sf <- permits_clean |>
  st_as_sf(coords = c("longitude", "latitude"), crs = CRS_GEO) |>
  st_transform(CRS_PROJ)
```

## Aggregate to tracts

```{r aggregate}
#| label: aggregate

# Spatial join: each permit -> tract GEOID
permits_tract <- st_join(permits_sf, chi_tracts[, "GEOID"], left = FALSE)

permit_summary <- permits_tract |>
  st_drop_geometry() |>
  group_by(GEOID) |>
  summarise(
    n_permits      = n(),
    total_value    = sum(reported_cost, na.rm = TRUE),
    .groups        = "drop"
  )

# Merge ACS + permits onto tract polygons
tract_data <- chi_tracts |>
  left_join(acs_raw, by = c("GEOID" = "geoid")) |>
  left_join(permit_summary, by = "GEOID") |>
  mutate(
    n_permits     = replace_na(n_permits, 0L),
    total_value   = replace_na(total_value, 0),
    value_per_hu  = ifelse(housing_units > 0, total_value / housing_units, NA_real_),
    log_value_hu  = log1p(value_per_hu)
  ) |>
  # Drop non-residential / empty tracts (e.g., O'Hare, industrial slivers)
  filter(housing_units >= 50, !is.na(total_pop), total_pop >= 100)
```

:::

# 3. Methodology

## 3.1 Method A — 2-Step Floating Catchment Area (2SFCA)

Standard 2SFCA (Luo & Wang 2003) computed against CTA "L" stations:

**Step 1** — for each station *j*, compute the station-to-population ratio
$R_j = S_j / \sum_{k \in N(j, d_0)} P_k$, where $S_j = 1$ (each station treated
as one unit of "rapid-transit supply"), $N(j, d_0)$ is the set of tract
centroids within $d_0 = 1{,}000$ m, and $P_k$ is tract *k* population.

**Step 2** — for each tract *i*, sum the ratios of all stations within
$d_0$ of its centroid: $A_i = \sum_{j \in N(i, d_0)} R_j$.

Euclidean distance is used as a documented simplification; street-network
distance would refine catchments, especially near rail lines and barriers, and is a logical extension.

```{r sfca}
#| label: sfca

# Tract centroids (projected)
tract_pts <- tract_data |>
  mutate(centroid = st_centroid(geometry)) |>
  st_drop_geometry() |>
  st_as_sf(sf_column_name = "centroid")

# Distance matrices (meters): tract <-> station
d_ts <- st_distance(tract_pts, l_stations) |> units::drop_units()
in_catch <- d_ts <= CATCHMENT_M  # logical matrix, rows = tracts, cols = stations

# --- Step 1: each station's served population, then R_j ---
pop_vec <- tract_pts$total_pop
station_pop <- colSums(in_catch * pop_vec, na.rm = TRUE)
R_j <- ifelse(station_pop > 0, 1 / station_pop, 0)

# --- Step 2: sum R_j over stations in each tract's catchment ---
A_i <- as.numeric(in_catch %*% R_j)

# Scale by 100,000 for human-readable magnitudes
tract_data$sfca_score <- A_i * 1e5
```

## 3.1.1 Catchment-Radius Sensitivity

Because Euclidean catchments are a modeling choice, the 2SFCA score is also
recomputed at 800 m and 1,200 m. The rank correlation against the headline
1,000 m score indicates whether the accessibility signal is stable to modest
radius changes.

```{r sfca-sensitivity}
#| label: sfca-sensitivity
#| results: asis

compute_sfca <- function(radius_m) {
  catch <- d_ts <= radius_m
  served_pop <- colSums(catch * pop_vec, na.rm = TRUE)
  ratios <- ifelse(served_pop > 0, 1 / served_pop, 0)
  as.numeric(catch %*% ratios) * 1e5
}

sfca_sensitivity <- tibble(radius_m = c(800, CATCHMENT_M, 1200)) |>
  mutate(score = map(radius_m, compute_sfca)) |>
  transmute(
    `Catchment radius` = paste0(radius_m, " m"),
    `Tracts with station access` = map_int(score, ~sum(.x > 0, na.rm = TRUE)),
    `Mean 2SFCA` = map_dbl(score, ~mean(.x, na.rm = TRUE)),
    `90th pct. 2SFCA` = map_dbl(score, ~quantile(.x, 0.9, na.rm = TRUE)),
    `Spearman rho vs. 1 km` = map_dbl(
      score,
      ~cor(.x, tract_data$sfca_score, method = "spearman", use = "complete.obs")
    )
  )

sfca_sensitivity |>
  gt() |>
  tab_header(title = md("**2SFCA Sensitivity to Catchment Radius**")) |>
  fmt_number(c(`Mean 2SFCA`, `90th pct. 2SFCA`), decimals = 2) |>
  fmt_number(`Spearman rho vs. 1 km`, decimals = 3) |>
  opt_row_striping()
```

## 3.2 Method B — Spatial Alpha Index + Local Moran's I

The **Spatial Alpha Index** combines the two signals on a common scale:

$$
\alpha_i = z(\text{SFCA}_i) - z(\log(1 + \text{value\_per\_hu}_i))
$$

Positive $\alpha_i$ = more accessibility than investment. We then test whether
high $\alpha$ values cluster spatially using Local Moran's I with queen
contiguity weights and 999 permutations.

```{r lisa}
#| label: lisa

# Build alpha index from z-scored components
tract_data <- tract_data |>
  mutate(
    z_sfca   = as.numeric(scale(sfca_score)),
    z_invest = as.numeric(scale(log_value_hu)),
    alpha    = z_sfca - z_invest
  )

# rgeoda needs an sf with no missing values on the focal variable
lisa_input <- tract_data |>
  filter(!is.na(alpha), is.finite(alpha))

w <- rgeoda::queen_weights(lisa_input)

set.seed(20260517)
lm_res <- rgeoda::local_moran(
  w, lisa_input["alpha"],
  permutations = 999
)

lisa_input$lisa_I       <- lisa_values(lm_res)
lisa_input$lisa_p       <- lisa_pvalues(lm_res)
lisa_input$lisa_cluster <- factor(
  lisa_labels(lm_res)[lisa_clusters(lm_res, cutoff = 0.05) + 1],
  levels = c("Not significant", "High-High", "Low-Low", "Low-High", "High-Low", "Undefined", "Isolated")
)

# Join LISA results back onto the full tract dataset
tract_data <- tract_data |>
  left_join(
    st_drop_geometry(lisa_input) |>
      select(GEOID, lisa_I, lisa_p, lisa_cluster),
    by = "GEOID"
  )
```

# 4. Results

## 4.1 Top-Alpha Tracts (publication table)

The table ranks tracts by alpha within the **High-High LISA cluster** — the
statistically significant under-invested + high-potential set. ACS profile
columns are included here so the target tracts can be interpreted as places,
not only as spatial-statistical outputs.

```{r gt-table}
#| label: gt-table
#| results: asis

top_alpha <- tract_data |>
  st_drop_geometry() |>
  filter(lisa_cluster == "High-High", lisa_p <= 0.05) |>
  arrange(desc(alpha)) |>
  slice_head(n = 15) |>
  transmute(
    GEOID,
    Population    = total_pop,
    `Housing units` = housing_units,
    `Median HH inc.` = med_hh_income,
    `% Renter`    = pct_renter,
    `% Non-white` = pct_nonwhite,
    `Permit $ / HU` = value_per_hu,
    `SFCA score`  = sfca_score,
    Alpha         = alpha,
    `LISA p`      = lisa_p
  )

top_alpha |>
  gt() |>
  tab_header(
    title    = md("**Chicago's Top 15 'Alpha' Tracts**"),
    subtitle = "High transit accessibility + low realized investment (HH cluster, p ≤ 0.05)"
  ) |>
  fmt_number(c(Population, `Housing units`), decimals = 0, use_seps = TRUE) |>
  fmt_currency(c(`Median HH inc.`, `Permit $ / HU`), decimals = 0) |>
  fmt_percent(c(`% Renter`, `% Non-white`), decimals = 1) |>
  fmt_number(c(`SFCA score`, Alpha), decimals = 2) |>
  fmt_number(`LISA p`, decimals = 3) |>
  data_color(
    columns = Alpha,
    fn      = scales::col_numeric(palette = "Greens", domain = NULL)
  ) |>
  data_color(
    columns = `Permit $ / HU`,
    fn      = scales::col_numeric(palette = "Reds", domain = NULL, reverse = TRUE)
  ) |>
  tab_source_note(md(
    "*Sources:* ACS 5-yr (2018-2022); Chicago Data Portal Building Permits 2019-2024; CTA L Stations."
  )) |>
  opt_row_striping() |>
  tab_options(table.font.size = px(13))
```

## 4.2 Citywide Summary

This table compares the target cluster against all other LISA categories. The
median income, renter share, and non-white share columns summarize the ACS
context behind the alpha signal.

```{r summary-gt}
#| label: summary-gt
#| results: asis

tract_data |>
  st_drop_geometry() |>
  filter(!is.na(lisa_cluster)) |>
  group_by(`LISA category` = lisa_cluster) |>
  summarise(
    `# tracts`        = n(),
    `Pop (median)`    = median(total_pop, na.rm = TRUE),
    `Med HH inc`      = median(med_hh_income, na.rm = TRUE),
    `% renter med`    = median(pct_renter, na.rm = TRUE),
    `% non-white med` = median(pct_nonwhite, na.rm = TRUE),
    `Permit $/HU med` = median(value_per_hu, na.rm = TRUE),
    `SFCA median`     = median(sfca_score, na.rm = TRUE),
    .groups = "drop"
  ) |>
  gt() |>
  tab_header(title = md("**Tract Profile by LISA Cluster**")) |>
  fmt_number(c(`Pop (median)`, `# tracts`), decimals = 0, use_seps = TRUE) |>
  fmt_currency(c(`Med HH inc`, `Permit $/HU med`), decimals = 0) |>
  fmt_percent(c(`% renter med`, `% non-white med`), decimals = 1) |>
  fmt_number(`SFCA median`, decimals = 2) |>
  opt_row_striping()
```

## 4.3 Interactive Maps

::: {.panel-tabset}

### Investment intensity

```{r map-invest}
#| label: map-invest
#| results: markup

geo <- tract_data |> st_transform(CRS_GEO)

pal_inv <- colorNumeric("YlOrRd", domain = geo$log_value_hu, na.color = "#dddddd")

leaflet(geo) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(
    fillColor = ~pal_inv(log_value_hu),
    weight    = 0.4, color = "white", fillOpacity = 0.85,
    label     = ~sprintf(
      "Tract %s<br>Permit $/HU: %s<br>Permits: %d",
      GEOID,
      dollar(value_per_hu, accuracy = 1),
      n_permits
    ) |> lapply(htmltools::HTML),
    highlightOptions = highlightOptions(weight = 2, color = "#222", bringToFront = TRUE)
  ) |>
  addLegend(
    "bottomright", pal = pal_inv, values = ~log_value_hu,
    title = "log(1 + $/HU)", opacity = 0.85
  )
```

### Transit accessibility (2SFCA)

```{r map-sfca}
#| label: map-sfca
#| results: markup

pal_sf <- colorNumeric("BuPu", domain = geo$sfca_score, na.color = "#dddddd")
stations_geo <- l_stations |> st_transform(CRS_GEO)

leaflet(geo) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(
    fillColor = ~pal_sf(sfca_score),
    weight    = 0.4, color = "white", fillOpacity = 0.85,
    label     = ~sprintf("Tract %s<br>SFCA: %.2f", GEOID, sfca_score) |>
      lapply(htmltools::HTML),
    group     = "Tracts"
  ) |>
  addCircleMarkers(
    data   = stations_geo,
    radius = 3, color = "#111", fillColor = "#fdae61",
    weight = 0.6, fillOpacity = 0.9, label = ~station_name,
    group  = "CTA L stations"
  ) |>
  addLayersControl(
    overlayGroups = c("Tracts", "CTA L stations"),
    options       = layersControlOptions(collapsed = FALSE)
  ) |>
  addLegend(
    "bottomright", pal = pal_sf, values = ~sfca_score,
    title = "2SFCA (×10⁵)", opacity = 0.85
  )
```

### LISA clusters of Alpha

```{r map-lisa}
#| label: map-lisa
#| results: markup

# Discrete color scheme for cluster categories
lisa_levels <- c("High-High", "Low-Low", "Low-High", "High-Low", "Not significant")
lisa_pal <- c(
  "High-High"       = "#d73027",
  "Low-Low"         = "#4575b4",
  "Low-High"        = "#abd9e9",
  "High-Low"        = "#fdae61",
  "Not significant" = "#e8e8e8"
)
geo$lisa_plot <- factor(
  ifelse(geo$lisa_p <= 0.05 & geo$lisa_cluster %in% lisa_levels,
         as.character(geo$lisa_cluster), "Not significant"),
  levels = lisa_levels
)
pal_lisa <- colorFactor(lisa_pal, levels = lisa_levels, na.color = "#e8e8e8")

leaflet(geo) |>
  addProviderTiles(providers$CartoDB.Positron) |>
  addPolygons(
    fillColor = ~pal_lisa(lisa_plot),
    weight    = 0.4, color = "white", fillOpacity = 0.85,
    label     = ~sprintf(
      "Tract %s<br>Alpha: %.2f<br>LISA: %s (p = %.3f)",
      GEOID, alpha, as.character(lisa_plot), lisa_p
    ) |> lapply(htmltools::HTML),
    group     = "LISA clusters"
  ) |>
  addCircleMarkers(
    data   = stations_geo,
    radius = 2.5, color = "#111", fillColor = "#ffffff",
    weight = 0.4, fillOpacity = 0.85,
    group  = "CTA L stations"
  ) |>
  addLayersControl(
    overlayGroups = c("LISA clusters", "CTA L stations"),
    options       = layersControlOptions(collapsed = FALSE)
  ) |>
  addLegend(
    "bottomright", pal = pal_lisa, values = lisa_levels,
    title = "LISA cluster (α index)", opacity = 0.9
  )
```

:::

## 4.4 Linked Browser (Top-Alpha Tracts)

Click a table row or map tract to highlight the same High-High alpha tract across both views.

```{r crosstalk}
#| label: crosstalk
#| results: markup

# Filter to the High-High cluster (target tracts), then build two parallel
# views: one sf for the map, one plain tibble for the table. Two SharedData
# objects with the SAME key + group keep them linked under crosstalk.
ct_sf <- tract_data |>
  filter(lisa_cluster == "High-High", lisa_p <= 0.05) |>
  st_transform(CRS_GEO) |>
  mutate(
    label = sprintf(
      "Tract %s | $/HU %s | SFCA %.2f | &alpha; %.2f",
      GEOID, dollar(value_per_hu, accuracy = 1), sfca_score, alpha
    )
  )

ct_tbl <- ct_sf |>
  st_drop_geometry() |>
  transmute(
    GEOID,
    Population      = total_pop,
    `Permit $/HU`   = round(value_per_hu, 0),
    `SFCA score`    = round(sfca_score, 2),
    Alpha           = round(alpha, 2),
    `LISA p`        = round(lisa_p, 3)
  )

shared_map <- SharedData$new(ct_sf,  key = ~GEOID, group = "alpha")
shared_tbl <- SharedData$new(ct_tbl, key = ~GEOID, group = "alpha")

bscols(
  widths = c(6, 6),
  leaflet(shared_map) |>
    addProviderTiles(providers$CartoDB.Positron) |>
    addPolygons(
      fillColor = "#d73027", weight = 0.6, color = "white",
      fillOpacity = 0.8, label = ~lapply(label, htmltools::HTML)
    ),
  datatable(
    shared_tbl,
    extensions = "Scroller",
    rownames   = FALSE,
    style      = "bootstrap",
    class      = "compact cell-border stripe nowrap",
    width      = "100%",
    options    = list(
      dom            = "tip",
      deferRender    = TRUE,
      scrollX        = TRUE,
      scrollY        = 420,
      scrollCollapse = TRUE,
      scroller       = TRUE,
      autoWidth      = TRUE,
      columnDefs     = list(
        list(width = "120px", targets = 0),
        list(width = "95px", targets = 1:5)
      )
    )
  )
)
```

# 5. Discussion

The High-High LISA cluster identifies 80 tracts where high alpha values are spatially concentrated: transit access is strong, but permit investment per housing unit remains comparatively low. The table in Section 4.1 lists the highest-alpha target tracts and includes ACS demographic profile columns. The summary in Section 4.2 shows that the High-High cluster has lower median household income than nonsignificant tracts, higher renter share (69.7%), higher non-white share (95.4%), and a much higher median 2SFCA score. This supports the impact-investing thesis, but it should be read as a screening signal for further neighborhood due diligence rather than as proof of investability on its own.

**Caveats.**

- 2SFCA uses Euclidean distance; a street-network refinement (e.g., via
  `sfnetworks` + OpenStreetMap) would tighten station catchments, especially
  where rail bisects neighborhoods.
- Building permits are a *flow* proxy for capital and undercount unpermitted
  rehab and institutional investment (LIHTC, opportunity-zone equity).
- The alpha index is descriptive, not causal. A natural extension is a
  **Spatial Lag Model** regressing investment on accessibility plus ACS
  controls, with the residuals reinterpreted as model-based alpha.

# Session info {.appendix}

```{r session}
#| label: session
#| results: markup
#| code-fold: true
sessionInfo()
```
